---
title: "Lab-8"
author: "Gowri"
  html:
    embed-resources: true
fig-height: 10
fig-width: 10
editor: visual
---

```{r}
library(parallel)
library(microbenchmark)

#This function generates an n x k dataset with all its entries drawn from a Poission distribution with mean lambda.

fun1 <- function(n = 100, k = 4, lambda = 4) {
  x <- NULL
  
  for (i in 1:n){
    x <- rbind(x, rpois(k, lambda))    
  }
  
  return(x)
}
#refrence - exampl- squae roots (cont. 3 ) shows vectorization principle 

fun1alt <- function(n = 100, k = 4, lambda = 4) {
  x <- matrix(rpois(n*k, lambda), nrow = n, ncol = k)
}


  



```

```{r}
#Benchmarking 
#show that funlalt generates a matrix with the same dimensions as fun1 and 
# that the values inside the 2 matricis folllow similar distriuitions

#Ysetseed! 
set.seed(1)
result1 <- fun1()
dim(result1) # dimension check 
mean(result1)
sd(result1)

set.seed(1)
result1alt <- fun1alt()
dim(result1alt) # dimension check 
mean(result1alt)
sd(result1alt)


#Then check the speed of the 2  functions wih rthe followinf codes 
microbenchmark::microbenchmark(
  fun1(),
  fun1alt()
)

```

2.  This function finds the maximum value of each column of a matrix.

```{r}

# Data Generating Process (10 x 10,000 matrix)
set.seed(1234)
x <- matrix(rnorm(1e4), nrow=10)

# Find each column's max value
fun2 <- function(x) {
  apply(x, 2, max)
}

fun2alt <- function(x) {
sapply(1:ncol(x), function(i) max(x[, i]))
}

#showing both teh functions are same 
 fun2(x)
 fun2alt(x)
 
 head(fun2(x))
 head(fun2alt(x))
 
 identical(fun2(x), fun2alt(x))
 
 #check the speed of. the 2 funnctions 
 microbenchmark::microbenchmark(
  fun2(x),
  fun2alt(x)
)

#Show that both functions return the same output for a given input matrix, x. Then check the speed of the two functions, as you did above.


```

Problem 3 :**Parallelization**

1.  This function implements a serial version of the bootstrap. Edit this function to parallelize the `lapply` loop, using whichever method you prefer. Rather than specifying the number of cores to use, use the number given by the `ncpus` argument, so that we can test it with different numbers of cores later.

    ```{r}


    # Use base R's built-in parallel package
    library(parallel)

    my_boot_serial <- function(dat, stat, R, ncpus = 1L) {
      n   <- nrow(dat)
      idx <- matrix(sample.int(n, n * R, TRUE), nrow = n, ncol = R)
      ans <- lapply(seq_len(R), function(i) stat(dat[idx[, i], , drop = FALSE]))
      ans <- do.call(rbind, ans)
      ans
    }

    my_boot <- function(dat, stat, R, ncpus = 1L) {
      n   <- nrow(dat)
      # Precompute indices for reproducibility across 1 vs multi-core
      idx <- matrix(sample.int(n, n * R, TRUE), nrow = n, ncol = R)

      if (ncpus > 1L) {
        cl <- parallel::makeCluster(ncpus)
        on.exit(parallel::stopCluster(cl), add = TRUE)

        # Make needed objects available on workers
        parallel::clusterExport(cl, varlist = c("dat", "idx", "stat"), envir = environment())

        ans <- parallel::parLapply(cl, seq_len(R), function(i) {
          stat(dat[idx[, i], , drop = FALSE])
        })
      } else {
        ans <- lapply(seq_len(R), function(i) {
          stat(dat[idx[, i], , drop = FALSE])
        })
      }

      ans <- do.call(rbind, ans)
      ans
    }

    ```

    ```{r}
    #2- Once you have a version of the my_boot() function that runs on multiple cores, check that it provides accurate results by comparing it to a parametric model:

    # Bootstrap of an OLS
    my_stat <- function(d) coef(lm(y ~ x, data=d))

    # DATA SIM
    set.seed(1)
    n <- 500; R <- 1e4

    x <- cbind(rnorm(n)); y <- x*5 + rnorm(n)

    # Checking if we get something similar as lm
    ans0 <- confint(lm(y~x))
    ans1 <- my_boot(dat = data.frame(x, y), my_stat, R = R, ncpus = 2L)

    # You should get something like this
    t(apply(ans1, 2, quantile, c(.025,.975)))
    ##                   2.5%      97.5%
    ## (Intercept) -0.1372435 0.05074397
    ## x            4.8680977 5.04539763
    ans0
    ##                  2.5 %     97.5 %
    ## (Intercept) -0.1379033 0.04797344
    ## x            4.8650100 5.04883353

    ```

    ```{r}
    system.time(my_boot(dat = data.frame(x, y), my_stat, R = 4000, ncpus = 1L))
    system.time(my_boot(dat = data.frame(x, y), my_stat, R = 4000, ncpus = 2L))
    ```
