---
title: "Lab-9"
author: "Gowri"
format: 
  html:
    embed-resources: true
fig-height: 10
fig-width: 10
editor: visual
---

```{r}
# install.packages(c("RSQLite", "DBI"))
library(RSQLite)
library(DBI)

# 1) In-memory SQLite database
con <- dbConnect(SQLite(), ":memory:")

# 2) Download tables (full, correct URLs)
actor    <- read.csv("https://raw.githubusercontent.com/ivanceras/sakila/master/csv-sakila-db/actor.csv")
rental   <- read.csv("https://raw.githubusercontent.com/ivanceras/sakila/master/csv-sakila-db/rental.csv")
customer <- read.csv("https://raw.githubusercontent.com/ivanceras/sakila/master/csv-sakila-db/customer.csv")
payment  <- read.csv("https://raw.githubusercontent.com/ivanceras/sakila/master/csv-sakila-db/payment_p2007_01.csv")

# 3) Copy into SQLite
dbWriteTable(con, "actor", actor)
dbWriteTable(con, "rental", rental)
dbWriteTable(con, "customer", customer)
dbWriteTable(con, "payment", payment)

# Quick checks
dbListTables(con)
dbGetQuery(con, "PRAGMA table_info(actor)")

```

# **Exercise 1**

Edit the code below to retrieve the actor ID, first name and last name for all actors using the `actor` table. Sort by last name and then by first name (note that the code chunk below is set up to run SQL code rather than R code).

```{sql connection=}
--| connection: con
SELECT actor_id, first_name, last_name
FROM actor
ORDER BY last_name, first_name;


```

# 

# **Exercise 2**

Retrieve the actor ID, first name, and last name for actors whose last name equals ‘WILLIAMS’ or ‘DAVIS’.

```{sql connection=}
--| connection: con
SELECT actor_id, first_name, last_name
FROM actor
WHERE last_name IN ('WILLIAMS', 'DAVIS');

```

# **Exercise 3**

Write a query against the `rental` table that returns the IDs of the customers who rented a film on July 5, 2005 (use the rental.rental_date column, and you can use the `date()` function to ignore the time component). Include a single row for each distinct customer ID.

```{sql connection=}
--| connection: con
SELECT DISTINCT customer_id
FROM rental
WHERE date(rental_date) = '2005-07-05';

```

# **Exercise 4**

## **Exercise 4.1**

Construct a query that retrieves all rows from the `payment` table where the amount is either 1.99, 7.99, 9.99.

```{sql connection=}
--| connection: con
SELECT *
FROM payment
WHERE amount IN (1.99, 7.99, 9.99);

```

## **Exercise 4.2**

Construct a query that retrieves all rows from the `payment` table where the amount is greater then 5.

```{sql connection=}
--| connection: con
SELECT *
FROM payment
WHERE amount > 5;

```

## **Exercise 4.2**

Construct a query that retrieves all rows from the `payment` table where the amount is greater then 5 and less then 8.

```{sql connection=}
--| connection: con
SELECT *
FROM payment
WHERE amount > 5 AND amount < 8;

```

# **Exercise 5**

Retrieve all the payment IDs and their amounts from the customers whose last name is ‘DAVIS’.

```{sql connection=}
--| connection: con
SELECT p.payment_id, p.amount
FROM payment AS p
INNER JOIN customer AS c
ON p.customer_id = c.customer_id
WHERE c.last_name = 'DAVIS';

```

## **Exercise 6.1**

Use `COUNT(*)` to count the number of rows in `rental`.

```{sql connection=}
--| connection: con
SELECT COUNT(*) AS total_rows
FROM rental;

```

## **Exercise 6.2**

Use `COUNT(*)` and `GROUP BY` to count the number of rentals for each `customer_id`.

```{sql connection=}
--| connection: con
SELECT customer_id, COUNT(*) AS rental_count
FROM rental
GROUP BY customer_id;

```

## **Exercise 6.3**

Repeat the previous query and sort by the count in descending order.

```{sql connection=}
--| connection: con
SELECT customer_id, COUNT(*) AS rental_count
FROM rental
GROUP BY customer_id
ORDER BY rental_count DESC;

```

## **Exercise 6.4**

Repeat the previous query but use `HAVING` to only keep the groups with 40 or more.

```{sql connection=}
--| connection: con
SELECT customer_id, COUNT(*) AS rental_count
FROM rental
GROUP BY customer_id
HAVING COUNT(*) >= 40
ORDER BY rental_count DESC;

```

### **Exercise 7**

Write a query that calculates a number of summary statistics for the payment table using `MAX`, `MIN`, `AVG` and

```{sql connection=}
--| connection: con
SELECT
  MAX(amount) AS max_amount,
  MIN(amount) AS min_amount,
  AVG(amount) AS avg_amount,
  SUM(amount) AS total_amount
FROM payment;

```

## **Exercise 7.1**

Modify the above query to do those calculations for each `customer_id`.

```{sql connection=}
--| connection: con
SELECT
  customer_id,
  MAX(amount) AS max_amount,
  MIN(amount) AS min_amount,
  AVG(amount) AS avg_amount,
  SUM(amount) AS total_amount
FROM payment
GROUP BY customer_id;

```

## **Exercise 7.2**

Modify the above query to only keep the `customer_id`s that have more then 5 payments.

```{sql connection=}
--| connection: con
SELECT
  customer_id,
  COUNT(*) AS num_payments,
  MAX(amount) AS max_amount,
  MIN(amount) AS min_amount,
  AVG(amount) AS avg_amount,
  SUM(amount) AS total_amount
FROM payment
GROUP BY customer_id
HAVING COUNT(*) > 5;

```

# **Cleanup**

Run the following chunk to disconnect from the connection.

```{r}
# clean up
dbDisconnect(con)

```
